### 2017-04-12 thesis/brainstorm.Rmd
adding brainstorm questions from week 6

---
title: "Brainstorm Questions for Melbourne Pedestrian Data"
output:
  html_document: default
  html_notebook: default
---
***

###Can we classify groups of sensor locations into types?###  
* Cluster analysis?  
* Types of areas? (restaurants and bars, clubs, shopping, train stations, etc..)  
    + peak traffic periods different?

###How does the pedestrian flow change with weather?###  
* Does rain result in less people in the city? Only on weekends and holidays?  
* Modelling based on past weather  
* Forecasting using weather forecast  
    + differences between business/office districts vs. leisure districts (shopping, cultural/tourist) - maybe lcassify these areas manually and see if we can confirm by cluster analysis?

###Is there a “maximum capacity” to how many people can move through the city in a given time frame? And if so, what affects this capacity?###

* At quantiles for given locations, say 95th percentile  
* Look at features of those locations to see what may affect flow (linked to clustering? maybe factor analysis?)  
    + Does this change with weather conditions? (similar to road traffic being affected by adverse weather potentially?)  
    + Any difference for areas with limited road traffic (Swanston and Bourke)

###Can we predict whether a day is a public holiday?###

* We expect that business areas to be quiet, and leisure areas to be just as busy/busier
    + maybe need to compare ratios between areas