### 2017-10-21 thesis_draft/index.Rmd

---
title: 'Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods'
degreetype: 'Bachelor of Economics Advanced with Honours'
author: 'Gavin Chin'
output: bookdown::pdf_book
site: bookdown::bookdown_site
bibliography: ["thesisrefs.bib"]
link-citations: yes
knit: "bookdown::render_book"
---

<!-- 
Edit these lines as appropriate.
The actual thesis content is in several Rmd files.

You'll need to edit the _bookdown.yml file to set the order in which you'd like them to appear.

If you have specific LaTeX packages to add, put them in monashthesis.tex.

You will need to ensure you have installed the knitr and bookdown packages for R.

You will also need LaTeX installed on your computer.
-->