### 2017-10-21 _book/thesis.tex
update

% This is a LaTeX thesis template for Monash University.
% to be used with Rmarkdown
% This template was produced by Rob Hyndman
% Version: 6 September 2016

\documentclass{monashthesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add any LaTeX packages and other preamble here if required
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Joe Bloggs}
\title{Thesis title}
\degrees{B.Sc. (Hons), University of Tangambalanga}
\def\degreetitle{Doctor of Philosophy}
% Add subject and keywords below
\hypersetup{
     %pdfsubject={The Subject},
     %pdfkeywords={Some Keywords},
     pdfauthor={Joe Bloggs},
     pdftitle={Thesis title},
     pdfproducer={Bookdown with LaTeX}
}

\bibliography{thesisrefs}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newtheorem{proposition}{Proposition}[chapter]
\theoremstyle{definition}
\newtheorem{example}{Example}[chapter]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[chapter]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\pagenumbering{roman}

\titlepage

{\sf\tighttoc}

\chapter*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my pet goldfish for \dots

\chapter*{Declaration}\label{declaration}
\addcontentsline{toc}{chapter}{Declaration}

I hereby declare that this thesis contains no material which has been
accepted for the award of any other degree or diploma in any university
or equivalent institution, and that, to the best of my knowledge and
belief, this thesis contains no material previously published or written
by another person, except where due reference is made in the text of the
thesis.

\vspace*{2cm}

\par

\authorname

\chapter*{Preface}\label{preface}
\addcontentsline{toc}{chapter}{Preface}

The material in Chapter \ref{ch:intro} has been submitted to the journal
\emph{Journal of Impossible Results} for possible publication.

The contribution in Chapter \ref{ch:litreview} of this thesis was
presented in the International Symposium on Nonsense held in Dublin,
Ireland, in July 2015.

\chapter*{Abstract}\label{abstract}
\addcontentsline{toc}{chapter}{Abstract}

This thesis is about \ldots{}

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

\chapter{Introduction}\label{ch:intro}

This is where you introduce the main ideas of your thesis, and an
overview of the context and background.

In a PhD, Chapter 2 would normally contain a literature review.
Typically, Chapters 3--5 would contain your own contributions. Think of
each of these as potential papers to be submitted to journals. Finally,
Chapter 6 provides some concluding remarks, discussion, ideas for future
research, and so on. Appendixes can contain additional material that
don't fit into any chapters, but that you want to put on record. For
example, additional tables, output, etc.

\section{Rmarkdown}\label{rmarkdown}

In this template, the rest of the chapter shows how to use Rmarkdown.
The big advantage of using Rmarkdown is that it allows you to include
your R code directly into your thesis, to ensure there are no errors in
copying and pasting, and that everything is reproducible. It also helps
you stay better organized.

For details on using \emph{R Markdown} see
\url{http://rmarkdown.rstudio.com}.

\section{Data}\label{data}

Included in this template is a file called \texttt{sales.csv}. This
contains quarterly data on Sales and Advertising budget for a small
company over the period 1981--2005. It also contains the GDP (gross
domestic product) over the same period. All series have been adjusted
for inflation. We can load in this data set using the following command:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sales <-}\StringTok{ }\KeywordTok{ts}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data/sales.csv"}\NormalTok{)[,-}\DecValTok{1}\NormalTok{], }\DataTypeTok{start=}\DecValTok{1981}\NormalTok{, }\DataTypeTok{frequency=}\DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Any data you use in your thesis can go into the data directory. The data
should be in exactly the format you obtained it. Do no editing or
manipulation of the data outside of R. Any data munging should be
scripted in R and form part of your thesis files (possibly hidden in the
output).

\section{Figures}\label{figures}

Figure \ref{fig:deaths} shows time plots of the data we just loaded.
Notice how figure captions and references work. Chunk names can be used
as figure labels with \texttt{fig:} prefixed. Never manually type figure
numbers, as they can change when you add or delete figures. This way,
the figure numbering is always correct.

\begin{figure}[htbp]
\centering
\includegraphics{thesis_files/figure-latex/deaths-1.pdf}
\caption{\label{fig:deaths}Quarterly sales, advertising and GDP data.}
\end{figure}

\section{Results from analyses}\label{results-from-analyses}

We can fit a dynamic regression model to the sales data.

If \(y_t\) denotes the sales in quarter \(t\), \(x_t\) denotes the
corresponding advertising budget and \(z_t\) denotes the GDP, then the
resulting model is:

\begin{equation}
  y_t - y_{t-4} = \beta (x_t-x_{t-4}) + \gamma (z_t-z_{t-4}) + \theta_1 \varepsilon_{t-1} + \Theta_1 \varepsilon_{t-4} + \varepsilon_t
\end{equation}

where \(\beta = 2.29\), \(\gamma = 0.99\), \(\theta_1 = 0.20\), and
\(\Theta_1 = -0.90\).

\section{Tables}\label{tables}

Let's assume future advertising spend and GDP are at the current levels.
Then forecasts for the next year are given in Table
\ref{tab:salesforecasts}.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lllll}
\toprule
Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
\midrule
1000.6 &  948.0 & 1053.3 & 920.2 & 1081.1 \\
1013.3 &  959.6 & 1066.9 & 931.2 & 1095.3 \\
1076.8 & 1023.1 & 1130.4 & 994.7 & 1158.8 \\
1003.3 &  949.6 & 1057.0 & 921.2 & 1085.4 \\
\bottomrule
\end{tabular}
\caption{Forecasts for the next year assuming Advertising budget and GDP are unchanged.}
\label{tab:salesforecasts}
\end{center}
\end{table}

Again, notice the use of labels and references to automatically generate
Table numbers. In this case, we need to generate the label ourselves.

The \texttt{knitLatex} package is useful for generating tables from R
output. Other packages can do similar things including the
\texttt{kable} function in \texttt{knitr} which is somewhat simpler but
you have less control over the result. If you use \texttt{knitLatex} to
generate tables, don't forget to include \texttt{results="asis"} in the
chunk settings.

\chapter{Literature Review}\label{ch:litreview}

This chapter contains a summary of the context in which your research is
set.

Imagine you are writing for your fellow PhD students. Topics that are
well-known to them do not have to be included here. But things that they
may not know about should be included.

Resist the temptation to discuss everything you've read in the last few
years. And you are not writing a textbook either. This chapter is meant
to provide the background necessary to understand the material in
subsequent chapters. Stick to that.

You will need to organize the literature review around themes, and
within each theme provide a story explaining the development of ideas to
date. In each theme, you should get to the point where your ideas will
fit in. But leave your ideas to later chapters. This way it is clear
what has been done beforehand, and what new contributions you are making
to the research field.

All citations should be done using markdown notation as shown below.
This way, your bibliography will be compiled automatically and
correctly.

\section{Exponential Smoothing}\label{sec:expsmooth}

Exponential smoothing was originally developed in the late 1950s
\autocites{Brown59}{Brown63}{Holt57}{Winters60}. Because of their
computational simplicity and interpretability, they became widely used
in practice.

Empirical studies by \textcite{MH79} and \textcite{Metal82} found little
difference in forecast accuracy between exponential smoothing and ARIMA
models. This made the family of exponential smoothing procedures an
attractive proposition \autocite[see][]{CKOS01}.

The methods were less popular in academic circles until \textcite{OKS97}
introduced a state space formulation of some of the methods, which was
extended in \textcite{HKSG02} to cover the full range of exponential
smoothing methods.

\appendix

\chapter{Additional stuff}\label{additional-stuff}

You might put some computer output here, or maybe additional tables.

Note that line 5 must appear before your first appendix. But other
appendices can just start like any other chapter.

\printbibliography[heading=bibintoc]

\end{document}