### 2017-10-23 thesis_draft/03-refs.Rmd
bib added as separate rmd, manually typed :(

---
knit: "bookdown::render_book"
---

# References {-}

City of Melbourne. (2017) City of Melbourne - Open Data Portal, data.melbourne.vic.gov.au

City of Melbourne. (2017) City Of Melbourne: Pedestrian Counting System, pedestrian.melbourne.vic.gov.au

David Kahle and Hadley Wickham. (2016) ggmap: Spatial visualization with ggplot2

Simon Garnier. (2017) The viridis color palettes

Lionel Henry and Hadley Wickham. (2017) purrr: Functional programming tools

Kirill Müller and Hadley Wickham. (2017) tibble: Simple data frames

Hyuk-Jae Roh, Satish Sharma, and Prasanta K. Sahu. (2016) Imputation of missing classified traffic data during winter season

Vitalie Spinu, Garrett Grolemund, and Hadley Wickham. (2016) lubridate: Make dealing with dates a little easier

Mario Pineda-Krch. (2010) Great-circle distance calculations in R, R-bloggers

Earo Wang. (2017) rwalkr: API to Melbourne Pedestrian Data

Hadley Wickham and Winston Chang. (2016) ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics

Hadley Wickham and Romain Francois and Lionel Henry and Kirill Müller. (2017) dplyr: A grammar of data manipulation

Jing-ting Zhong and Shuai Ling. (2015) Key factors of k-nearest neighbours nonparametric regression in short-time traffic flow forecasting, Proceedings of the 21st International Conference on Industrial Engineering and Engineering Management 2014, 9–12

Ming Zhong, Satish Sharma, and Zhaobin Liu. (2005) Assessing robustness of imputation models based on data from different jurisdictions: Examples of Alberta and Saskatchewan, Canada, Transportation Research Record: Journal of the Transportation Research Board 1917, 116–126

Nina Zumel. (2014) Trimming the fat from glm() models in R, Win-Vector LLC